Data Analysis and Modelling of Teen Vaccination Surveys

### 2024-03-15 code/analysis-code/analysis1.qmd
Analysis for multicollinearity

I consulted ChatGPT to ask how to perform a VIF analysis in R and was recommended the packages 'car' and 'carData' using the vif() function.
I will start by opening the libraries necessary.
```{r}
library(ggplot2)
library(car) # to perform VIF analysis
library(carData) # to look at
```

Next I will produce a residual plot
```{r}
model <- glm(glm(P_UTDHPV ~ AGE + SEX + STATE + INS_STAT2_I + INS_BREAK_I + INCPOV1 + RACEETHK + EDUC1 + LANGUAGE + MOBIL_1 + FACILITY, data = mydata2, family = binomial)
# Residual plot
residualPlot <- function(model) {
  ggplot(data = mydata2(fit = predict(model), residuals = resid(model)),
         aes(x = fit, y = residuals)) +
    geom_point() +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    xlab("Fitted values") +
    ylab("Residuals") +
    ggtitle("Residual Plot")
}

# VIF diagnostics
vifDiagnostics <- function(model) {
  vif_values <- car::vif(model)
  print(vif_values)
}

# Create residual plot
residual_plot <- residualPlot(fitall)
print(residual_plot)

# Calculate VIFs
vif_diagnostics <- vifDiagnostics(model)
print(vif_diagnostics)

```
```

>>>>

I consulted ChatGPT to ask how to perform a VIF analysis in R and was recommended the packages 'car' and 'carData' using the vif() function.
I will start by opening the libraries necessary.
```{r}
library(ggplot2)
library(car) # to perform VIF analysis
library(carData)
```

Next I will produce a residual plot. I asked ChatGPT how to use ggplot to produce a residual plot for a logistic regression function and was given the base code, which I modified to fit my analysis.
```{r}
# First, I will define the model given to me by the step wise comparison
model <- glm(P_UTDHPV ~ AGE + SEX + STATE + INS_STAT2_I + INS_BREAK_I + INCPOV1 + RACEETHK + EDUC1 + LANGUAGE + MOBIL_1 + FACILITY, data = mydata2, family = binomial)

# Define function for residual plot of logistic regression
residualPlotLogit <- function(model) {
  fitted_prob <- predict(model, type = "response") # Predicted probabilities
  residuals <- resid(model, type = "response") # Calculate residuals
  plot_data <- data.frame(fit = fitted_prob, residuals = residuals) # Create data frame

  # Plot
  ggplot(data = plot_data,
         aes(x = fit, y = residuals)) +
    geom_point() +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    xlab("Fitted probabilities") +
    ylab("Residuals") +
    ggtitle("Residual Plot of Proposed Stepwise Model")
}

# Create the residual plot
residual_plot_logit <- residualPlotLogit(fitall)
print(residual_plot_logit)

```
The residual plot shows clear collinearity that is not represented by the current model. To find the predictors with collinnearity we will use a correlation matrix. This is because the VIF analysis will not run when multicollinearity is too high.
```{r}
factor_data <- model.matrix(~ STATE + FACILITY + RACEETHK + AGE + EDUC1 + INS_BREAK_I + INCPOV1 + SEX + LANGUAGE + MOBIL_1 + INS_STAT2_I, data = mydata2)

# Compute correlation matrix
correlation_matrix <- cor(factor_data)

# Find indices of correlation matrix where values are 1 or 0
indices <- which(!is.na(correlation_matrix) & correlation_matrix == 1 | correlation_matrix == 0, arr.ind = TRUE)

# Extract variable names corresponding to these indices
variable_pairs <- rownames(correlation_matrix)[indices[,1]]
variable_pairs <- cbind(variable_pairs, colnames(correlation_matrix)[indices[,2]])

# Print variable pairs with correlation values of 1 or 0
print(variable_pairs)

```
This shows that many levels from each of the variables show multicollineaity. This might be addressed by using a different model for analysis.
